---
title: "hw1_mra2209"
author: "Maya Arnott"
date: "2025-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(tidyverse)
library(moderndive)
```

## Loading data

```{r}
data("early_january_weather")
```

The data set contains hourly meteorological data for the airports LGA, JFK, and EWR for the month of January 2013. It contains the following variables:  

  - **origin**, indicating the weather station.  
  
  - **year**, month, day, hour, indicating the time of recording.  
  
  - **temp**, dewp, indicating the temperature and dewpoint (in F).  
  
  - **humid**, indicating the relative humidity.  
  
  - **wind_dir**, **wind_speed**, **wind_gust**, indicating the wind            direction (in degrees), speed and gust speed (in mph).  
  
  - **precip**, indicating the precipitation (in inches). 
  
  - **pressure**, indicating the sea level pressure (in millibars).
  
  - **visib**, indicating the visibility (in miles).  
  
  - **time_hour**, indicating the date and hour of the recording.
  
The data set has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. The mean temperature is `r early_january_weather |> summarize(mean_temp = mean(temp, na.rm = TRUE)) |> pull(mean_temp) |> round(2)` degrees Fahrenheit. 


## Creating a Scatter Plot
```{r}

weather_plot <- early_january_weather |> 
  ggplot( aes ( x = time_hour, y = temp, color = humid )) +
  geom_point(alpha = 0.7) + 
  labs(
      title = "Temperature over Time, Colored by Humidity",
      x = "Time (date & hour)",
      y = "Temperature (°F)",
      color = "Humidity"
  ) + 
  theme_minimal() + 
  scale_color_gradient(low = "blue", high = "red") + 
  theme(plot.title = element_text( 
    hjust = 0.5,
    face = "bold",
    size = 14
    )
  )

weather_plot
```


## Saving the plot using ggsave()

```{r}

ggsave("weather_plot.png", plot = weather_plot, width = 8, height = 5 )

```

