---
title: "p8105_hw1_mra2209"
author: "Maya Arnott"
date: "2025-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(tidyverse)
library(tibble)
library(moderndive)
```

## Loading data

```{r}
data("early_january_weather")
```

The data set contains hourly meteorological data for the airports LGA, JFK, and EWR for the month of January 2013. It contains the following variables:  

  - **origin**, indicating the weather station.  
  
  - **year**, month, day, hour, indicating the time of recording.  
  
  - **temp**, dewp, indicating the temperature and dewpoint (in F).  
  
  - **humid**, indicating the relative humidity.  
  
  - **wind_dir**, **wind_speed**, **wind_gust**, indicating the wind            direction (in degrees), speed and gust speed (in mph).  
  
  - **precip**, indicating the precipitation (in inches). 
  
  - **pressure**, indicating the sea level pressure (in millibars).
  
  - **visib**, indicating the visibility (in miles).  
  
  - **time_hour**, indicating the date and hour of the recording.
  
The data set has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. The mean temperature is `r early_january_weather |> summarize(mean_temp = mean(temp, na.rm = TRUE)) |> pull(mean_temp) |> round(2)` degrees Fahrenheit. 


## Creating a Scatter Plot
```{r}

weather_plot <- early_january_weather |> 
  ggplot( aes ( x = time_hour, y = temp, color = humid )) +
  geom_point(alpha = 0.7) + 
  labs(
      title = "Temperature over Time, Colored by Humidity",
      x = "Time (date & hour)",
      y = "Temperature (°F)",
      color = "Humidity"
  ) + 
  scale_x_datetime(
    date_breaks = "48 hours",
    date_labels = "%b %d\n%H:%M"
  ) +
  theme_minimal() + 
  scale_color_gradient(low = "blue", high = "red") + 
  theme(plot.title = element_text( 
    hjust = 0.5,
    face = "bold",
    size = 14
    )
  )

weather_plot
```

In the Temperature over Time plot above, we can see how the temperature around the LGA, JFK, and EWR airports changes over time during early January 2013. There appears to be a general day-night cycle in the temperatures. The humidity color gradient shows that higher humidity is usually associated with warmer temperatures and that low humidity is usually associated with colder times. There may also be periodic shifts due to weather fronts or the time of day.
r

## Saving the plot using ggsave()

```{r}

ggsave("weather_plot.png", plot = weather_plot, width = 8, height = 5 )

```

## Creating a data frame for Problem 2

```{r}

my_df <- tibble (
  
  sample_nums = rnorm(10),
  
  is_positive = sample_nums > 0,
  
  char_vec = sample(letters, 10),
  
  factor_vec = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))

  )

```




